---
title: "Soluble IL-2R after ICI — Analysis"
subtitle: "Who had a soluble IL-2 receptor drawn after an immune checkpoint inhibitor?"
author: "Tianqi Ouyang"
date: "`r Sys.Date()`"
execute:
  eval: true
  echo: true
  warning: false
  message: false
---

::: {.callout-note}
## About this page

This page answers the four cohort questions for the soluble IL-2 receptor (sIL2R)
project. Every number is an **aggregate** read from `cache/summary.rds`, which
`R/build.R` writes locally from the RPDR extract and the ICI data pull. Counts from
1 to 10 are shown as `<11`. Patient-level detail (EMPI, dates) stays in the local
workbook `output/sIL2R_ICI_patient_level.xlsx` and is not part of this site.
:::

```{r}
#| label: setup
#| include: false
suppressPackageStartupMessages({ library(data.table); library(knitr) })
source("R/functions.R")
have <- file.exists("cache/summary.rds")
if (have) S <- readRDS("cache/summary.rds")
```

```{r}
#| label: not-built
#| echo: false
#| output: asis
if (!have) cat("::: {.callout-warning}\n## Pipeline has not been run\n",
               "Run `Rscript R/build.R` locally, then re-render.\n:::\n")
```

```{r}
#| label: not-built-exit
#| include: false
if (!have) knitr::knit_exit()
a   <- S$answers
N0  <- a[["post_cutoff_patients"]]
N1  <- a[["q1_ici_before_draw"]]
N2  <- a[["q2_ge2_after_ici"]]
N3  <- a[["q3_not_in_ici_cohort"]]
NB  <- a[["in_ici_not_before"]]
pct <- function(n, d) formatC(100 * n / d, format = "f", digits = 1)
## the draws-vs-days gap is itself a patient count: show it only when safe
dq2 <- a[["q2_ge2_after_ici_draws"]] - N2
q2_note <- if (dq2 == 0) " (the same if two draws on one day are counted separately)" else
  if (dq2 >= SMALL_CELL) sprintf(" (%s if two draws on one day are counted separately)",
                                 fmt_n(a[["q2_ge2_after_ici_draws"]])) else
  " (fewer than 11 more if two draws on one day are counted separately)"
```

# 1 The requests, answered

```{r}
#| label: tbl-answers
#| echo: false
data.table(
  Request = c("RPDR patients with a numeric soluble IL2R Result on/after 2015-01-01 (all-pre-2015 excluded)",
              "... who had gotten an ICI (any dose) before an IL2R level was drawn",
              "... of those, with at least 2 IL2R measurements after the ICI",
              "... NOT found in the ICI cohort at all (chart-review candidates)"),
  N   = fmt_n(c(N0, N1, N2, N3)),
  `%` = c("", paste0(pct(N1, N0), "% of ", fmt_n(N0)),
          paste0(pct(N2, N1), "% of ", fmt_n(N1)),
          paste0(pct(N3, N0), "% of ", fmt_n(N0)))) |>
  kable(caption = "Primary definitions (Section 2). 'After 2015' = drawn on or after 2015-01-01; 'before' = ICI start date strictly earlier than the IL2R draw date; a measurement = one draw date with a numeric value in the RPDR Lab `Result` field.")
```

The remaining `r fmt_n(NB)` patients are in the ICI cohort but started ICI on or
after the date of every post-2015 IL2R draw, so
`r fmt_n(N0)` = `r fmt_n(N1)` + `r fmt_n(NB)` + `r fmt_n(N3)`.

**Not counted above:** `r fmt_n(S$excluded$n[1])` patients are in the RPDR patient
set but have **no IL2R row in the RPDR Lab file**, so there is no draw date and they
cannot be placed before or after ICI. They split as follows.

```{r}
#| label: tbl-nolab
#| echo: false
S$nolab[, .(`IL2R order, no result row` = fifelse(in_ici_file, "In the ICI pull (request 1 status unknown)",
                                                  "Not in the ICI pull (add to request 3 review)"),
            Patients = fmt_n(patients),
            `ICI in RPDR Med` = fmt_n_of(rpdr_ici_any, patients),
            `Malignancy code` = fmt_n_of(cancer_dx_any, patients))] |>
  kable(caption = "Patients without a draw date. Dating them needs the RPDR procedures (Prc) file for the same patient set, or chart review; they are listed in the local workbook (sheet no_IL2R_result_row).")
```

# 2 Definitions as implemented

The code below is the **executed** definition, pulled verbatim from
`R/functions.R` at render time — click **Show code** on any block.

## 2.1 Soluble IL2R draws (RPDR Lab)

An IL2R draw is any RPDR Lab row with `Group_Id` in the three sIL2R groups; the
draw date is the date part of `Seq_Date_Time`. A row counts as an IL2R **measurement
only when its `Result` field starts with a number** (e.g. `1308` or `1234.5 pg/mL H`).
Rows whose `Result` is blank or text only (e.g. *SEE MANUAL REPORT*,
*CREDITED: NO SPECIMEN RECEIVED*) are not measurements. The free-text report field
`Result_Text` is not used.

```{r}
#| label: def-il2r
#| eval: false
#| code: !expr src_of(c("IL2R_GROUPS", "CUTOFF_DATE", "NUM_REGEX", "STATUS_LEVELS", "classify_il2r_result"))
```

## 2.2 ICI start date (ICI data pull)

On read, `EMPI MRN` (column G) is renamed to `EMPI` so it keys directly against the
RPDR files, and `SCHEDULEDSTARTDTS` (raw `YYYY-MM-DDThh:mm:ssZ`) becomes a date
(`YYYY-MM-DD`). Every row in the pull is a MAR *New Bag* administration, so every
row is a dose given. The ICI start date is the earliest administration date, not
counting blinded *"X OR PLACEBO"* trial products (the patient may have had placebo).

```{r}
#| label: def-ici
#| eval: false
#| code: !expr src_of(c("read_ici", "parse_ici_date", "is_blinded"))
```

## 2.3 Patient classification

```{r}
#| label: def-cohort
#| eval: false
#| code: !expr src_of(c("derive_cohort", "answer_counts"))
```

# 3 Cohort flow

```{r}
#| label: tbl-flow
#| echo: false
S$flow[, .(Step = step, N = fmt_n(n))] |>
  kable(caption = "From the RPDR patient set to the post-2015 sIL2R cohort.")
```

```{r}
#| label: tbl-excluded
#| echo: false
rbind(S$excluded[, .(`Not in the post-2015 cohort because` = reason, N = fmt_n(n))],
      S$excluded_sub[, .(`Not in the post-2015 cohort because` = reason, N = fmt_parts(n))]) |>
  kable(caption = "Patients dropped at each step of the cohort flow above; the last two rows split the second row. A row whose `Result` is blank or text only (e.g. SEE MANUAL REPORT, CREDITED: ...) is not a measurement.")
```

# 4 Request 1 — ICI before the IL2R draw

```{r}
#| label: tbl-groups
#| echo: false
g <- S$groups
b2_note <- if (!GROUP_LEVELS[3] %in% as.character(g$group))
  " Group B2 (only a blinded or-placebo trial product) is empty in this cohort." else ""
data.table(Group = as.character(g$group), N = fmt_parts(g$N),
           `%` = pct_shown(fmt_parts(g$N), g$N, sum(g$N))) |>
  kable(caption = sprintf("Every patient in the post-2015 cohort (N = %s) falls in exactly one group.%s", fmt_n(N0), b2_note))
```

## 4.1 Time from ICI start to the first post-ICI draw

```{r}
#| label: tbl-time
#| echo: false
S$A_time_bins[, .(`ICI start to first IL2R draw` = as.character(bin),
                  N = fmt_parts(N), `%` = pct_shown(fmt_parts(N), N, sum(N)))] |>
  kable(caption = sprintf("Group A (N = %s). Median %s days (IQR %s-%s).",
                          fmt_n(S$A_days_to_first$n), S$A_days_to_first$median,
                          round(S$A_days_to_first$p25), round(S$A_days_to_first$p75)))
```

## 4.2 First ICI regimen

```{r}
#| label: tbl-first-drug
#| echo: false
fd <- collapse_small(S$A_first_regimen, "ici_first_regimen")
fd[, .(`ICI given on the start date` = ici_first_regimen, N = fmt_parts(N),
       `%` = pct_shown(fmt_parts(N), N, sum(N)))] |>
  kable(caption = "Group A. Every ICI product given on the ICI start date, so combination starts (e.g. ipilimumab + nivolumab) show as one regimen; each product is labelled by the first ICI or fixed-dose co-formulation named in its description. Regimens with fewer than 11 patients are combined as 'Other'.")
```

## 4.3 Calendar year

```{r}
#| label: tbl-year
#| echo: false
yr <- copy(S$groups_by_first_year)
yr[, period := fifelse(year <= 2020, "2015-2020", as.character(year))]
cols <- setdiff(names(yr), c("year", "period"))
yr <- yr[, lapply(.SD, sum), by = period, .SDcols = cols]
setnames(yr, cols, sub("^([A-Z0-9]+)\\..*$", "Group \\1", cols))
yr[, (2:ncol(yr)) := lapply(.SD, fmt_n), .SDcols = 2:ncol(yr)]
setnames(yr, "period", "Year of first post-2015 IL2R draw")
gA <- S$groups_by_first_year[[GROUP_LEVELS[1]]]
a_pre2021 <- if (is.null(gA)) 0 else sum(gA[S$groups_by_first_year$year <= 2020])
kable(yr, caption = paste0(
  "IL2R testing with a numeric Result is concentrated from 2021, when the Epic-coded assay (el:5200007493) begins; Group A is therefore ",
  if (a_pre2021 == 0) "entirely" else "mostly", " in 2021 onward."))
```

# 5 Request 2 — at least two measurements after ICI

**`r fmt_n(N2)`** of the `r fmt_n(N1)` patients in Group A had IL2R measured on at
least two different days after the ICI start date`r q2_note`. Group A contributes
`r fmt_n(S$A_measurements_total)` post-ICI measurement days in total.

```{r}
#| label: tbl-n-after
#| echo: false
na <- copy(S$A_n_after)
na[, k := fifelse(n_after %in% c("4", "5+"), "4 or more", n_after)]
na <- na[, .(N = sum(N)), by = k]
na[, .(`Post-ICI IL2R measurement days` = k, N = fmt_parts(N), `%` = pct_shown(fmt_parts(N), N, sum(N)))] |>
  kable(caption = sprintf("Group A (N = %s). 'At least 2' = the sum of the last three rows.", fmt_n(N1)))
```

# 6 Request 3 — not found in the ICI cohort

**`r fmt_n(N3)`** of the `r fmt_n(N0)` post-2015 patients (`r pct(N3, N0)`%) have no
record at all in the ICI data pull. Before deciding on chart review, two independent
RPDR sources were checked for the same patients.

```{r}
#| label: tbl-med
#| echo: false
S$med_by_group[, .(Group = as.character(group), Patients = fmt_n(patients),
      `ICI in RPDR Med (any time)` = fmt_n_of(rpdr_ici_any, patients),
      `ICI in RPDR Med before last IL2R draw` = fmt_n_of(rpdr_ici_before_last_draw, patients))] |>
  kable(caption = "Cross-check against the RPDR Med file: any ICI generic, sponsor code or brand name. 'suppressed' = the complement within the group is below 11.")
```

The RPDR medication file finds ICI in most of Group A, which shows it captures ICI
exposure well — yet it finds ICI in almost none of Group C. As a second triage signal,
every Group C patient was checked in the RPDR Dia file for a malignancy code
(ICD-10 C00–C96 including C4A/C7A/C7B; ICD-9 140–209.3) and, failing that, a
cancer-history or chemotherapy/immunotherapy code (Z85, Z51.1x, Z92.21; V10, V58.1x).
All `r fmt_n(S$dx_by_group[group == GROUP_LEVELS[1], cancer_dx_any])` Group A
patients have a malignancy code.

```{r}
#| label: tbl-triage
#| echo: false
S$C_triage[, .(`Chart-review tier (Group C)` = tier, N = fmt_parts(N), `%` = pct_shown(fmt_parts(N), N, sum(N)))] |>
  kable(caption = sprintf("Group C (N = %s), hierarchical: each patient is placed in the first tier that applies.", fmt_n(N3)))
```

```{r}
#| label: triage-n
#| include: false
tr <- setNames(S$C_triage$N, substr(S$C_triage$tier, 1, 1))
medA <- S$med_by_group[group == GROUP_LEVELS[1]]
```

::: {.callout-tip}
## Suggested chart-review scope

The RPDR medication file records an ICI for `r pct(medA$rpdr_ici_any, medA$patients)`%
of Group A, so it is a reasonable second net — and it finds ICI for only
`r fmt_n(tr[["1"]])` Group C patients. Review those first. The
`r fmt_n(tr[["2"]])` tier-2 patients (a malignancy code on or before the last IL2R
draw) are where ICI given outside MGB, or before Epic, is still plausible. Tier-5
patients (`r fmt_n(tr[["5"]])`) have no malignancy, cancer-history or chemotherapy
code in RPDR, so ICI exposure is unlikely — though RPDR can miss cancer care given
entirely outside MGB — and a small random sample of them would confirm that before
skipping the rest.

Also worth a look: `r fmt_n(S$B_rpdr_earlier)` Group B patients have an ICI in the
RPDR medication file **before** their last IL2R draw, even though their first ICI in
the pull comes later. They may belong in Group A.

Group C tiers are in the local workbook (sheet `not_in_ICI_cohort`, column
`review_tier`); the flagged Group B patients are in sheet `ICI_cohort_RPDR_ICI_earlier`.
:::

# 7 Sensitivity analyses

```{r}
#| label: tbl-sens
#| echo: false
sv   <- S$sensitivity
cols <- c("post_cutoff_patients", "q1_ici_before_draw", "q2_ge2_after_ici", "q3_not_in_ici_cohort")
out  <- copy(sv[, c("analysis", cols), with = FALSE])
for (cc in cols) out[, (cc) := c(fmt_n(sv[[cc]][1]), fmt_delta(sv[[cc]][-1], sv[[cc]][1]))]
## cohort = R1 + B + R3, so (change in cohort) - (change in R3) is the change in
## ICI-pull patients; when that is 1-10 the cohort change is also hidden.
d  <- function(cc) sv[[cc]][-1] - sv[[cc]][1]
d_in <- d("post_cutoff_patients") - d("q3_not_in_ici_cohort")
hide <- c(FALSE, abs(d_in) >= 1 & abs(d_in) < SMALL_CELL)
out[hide, post_cutoff_patients := "suppressed"]
setnames(out, c("Analysis", "Post-2015 cohort", "R1: ICI before draw",
                "R2: >= 2 days after ICI", "R3: not in ICI cohort"))
m12   <- max(abs(c(d("q1_ici_before_draw"), d("q2_ge2_after_ici"))))
r3max <- sv$analysis[-1][which.max(abs(d("q3_not_in_ici_cohort")))]
kable(out, caption = paste0(
  "Primary row = counts; other rows = change from primary. Changes of 1-10 are shown as +/-<11, ",
  "and a cohort change is 'suppressed' when it would reveal one by subtraction. R1 and R2 move by ",
  if (m12 < SMALL_CELL) "fewer than 11 patients" else paste("up to", fmt_n(m12), "patients"),
  " under every alternative definition; R3 moves most under '", r3max, "'."))
```

# 8 Caveats

- **ICI coverage starts `r format(S$ici_date_range[1], "%B %Y")`.** The ICI pull
  holds Epic MAR administrations from `r format(S$ici_date_range[1], "%B %Y")` to
  `r format(S$ici_date_range[2], "%B %Y")`. ICI given before Epic, or outside
  MGB, is not in it — which is what the RPDR medication check in Section 6 is for.
- **IL2R results in the Lab file are uneven over time.** The Epic-coded assay
  (el:5200007493) starts in `r S$il2_first_month_epic`; few results exist for
  2018–2020, and the latest result in the extract is from `r S$il2_last_month`.
  IL2R tests from those gaps may belong to the patients with no IL2R row in the
  Lab file (@tbl-nolab).
- **Date resolution.** ICI start is a calendar date, so an IL2R drawn on the ICI
  start day is counted as *not* after ICI (`r fmt_n(S$same_day_A + S$same_day_B)`
  patients affected; see the 'Same-day ICI counts as prior' row of @tbl-sens).
- **Units differ by assay** (pg/mL for the Epic-coded assay, U/mL for the older
  mcsq/dcsqe assays). This does not affect any count here, but values must be
  harmonised before they are compared.

```{r}
#| label: session
#| echo: false
#| output: asis
cat("\n---\n\n*Aggregate results only; cells below 11 suppressed. Built from `R/build.R` on ",
    format(S$built, "%Y-%m-%d"), ".*\n", sep = "")
```
