---
title: "Soluble IL-2R × ICI — Data Management"
subtitle: "Inputs, variable changes and linkage QC"
author: "Tianqi Ouyang"
date: "`r Sys.Date()`"
---

::: {.callout-note}
## About this page

All patient-level processing happens locally in `R/build.R` (run with
`Rscript R/build.R`). This page shows the code and the **aggregate** QC tables the
build writes to `cache/summary.rds`. Counts from 1 to 10 are shown as `<11`.
:::

```{r}
#| label: setup
#| include: false
suppressPackageStartupMessages({ library(data.table); library(knitr) })
source("R/functions.R")
have <- file.exists("cache/summary.rds")
if (have) S <- readRDS("cache/summary.rds")
```

```{r}
#| label: not-built
#| echo: false
#| output: asis
if (!have) cat("::: {.callout-warning}\n## Pipeline has not been run\n",
               "Run `Rscript R/build.R` locally, then re-render.\n:::\n")
```

```{r}
#| label: exit
#| include: false
if (!have) knitr::knit_exit()
```

# 1 Data sources

```{r}
#| label: tbl-sources
#| echo: false
data.table(
  Source = c("RPDR Mrn", "RPDR Lab", "RPDR Med", "RPDR Dia", "ICI data pull"),
  File   = c(sub("\\{.*", "Mrn.txt", S$files$rpdr), sub("\\{.*", "Lab.txt", S$files$rpdr),
             sub("\\{.*", "Med.txt", S$files$rpdr), sub("\\{.*", "Dia.txt", S$files$rpdr),
             S$files$ici),
  Use    = c("Patient set (one row per patient)",
             "Soluble IL2R results (plus creatinine, not used here)",
             "Independent check for ICI exposure",
             "Malignancy codes, for chart-review triage",
             "ICI administrations (Epic MAR 'New Bag')"),
  Rows   = c(fmt_n(S$flow$n[1]), fmt_n(S$lab_rows), fmt_n(S$med_rows), "-", fmt_n(S$ici_rows)),
  Patients = c(fmt_n(S$flow$n[1]), fmt_n(S$lab_patients), "-", "-", fmt_n(S$ici_patients))) |>
  kable(caption = "Inputs. The RPDR extract (IRB 2022P001091) was pulled 2026-08-11; the ICI pull is dated 2026-07-17. Neither is in this repository.")
```

# 2 ICI data pull

## 2.1 Variable changes

Two changes are made when the file is read, so every downstream object uses the
RPDR conventions:

| Raw | Cleaned | Why |
|---|---|---|
| `EMPI MRN` (column G) | `EMPI` | Same name as the RPDR key; joins need no renaming |
| `SCHEDULEDSTARTDTS` = `YYYY-MM-DDThh:mm:ssZ` | `SCHEDULEDSTARTDTS` = `YYYY-MM-DD` (Date) | ICI timing is compared at the day level |

```{r}
#| label: code-read-ici
#| eval: false
#| code: !expr src_of(c("read_ici", "parse_ici_date"))
```

```{r}
#| label: tbl-ici-format
#| echo: false
data.table(
  Check = c("`EMPI` values that are 9 digits (ICI pull)",
            "`EMPI` values that are 9 digits (RPDR Lab, IL2R rows)",
            "`SCHEDULEDSTARTDTS` values that failed to parse",
            "Earliest `SCHEDULEDSTARTDTS` (month)", "Latest `SCHEDULEDSTARTDTS` (month)"),
  Result = c(sprintf("%.1f%%", 100 * S$ici_empi_9digit),
             sprintf("%.1f%%", 100 * S$rpdr_empi_9digit),
             fmt_n(S$ici_date_na),
             format(S$ici_date_range[1], "%Y-%m"), format(S$ici_date_range[2], "%Y-%m"))) |>
  kable(caption = "Format checks. Both files carry EMPI as a 9-digit string, so they are joined as character without padding.")
```

# 3 RPDR soluble IL2R labs

## 3.1 Records by year

```{r}
#| label: tbl-il2-year
#| echo: false
y <- copy(S$il2_by_year)
y[, period := fcase(year <= 2014, "2005-2014", year %between% c(2018, 2020), "2018-2020",
                    default = as.character(year))]
y <- y[, .(records = sum(records)), by = period]
y[, .(`Draw year` = period, `IL2R records` = fmt_n(records))] |>
  kable(caption = "IL2R rows by year (all rows, numeric Result or not). Results before 2015 are excluded from the cohort; the Epic-coded assay begins in 2021.")
```

# 4 Linkage

```{r}
#| label: tbl-linkage
#| echo: false
data.table(
  Check = c("IL2R patients (Lab) not in the RPDR patient set",
            "Post-2015 IL2R patients found in the ICI pull",
            "Patient-days with 2 or more numeric IL2R Results"),
  N = fmt_n(c(S$il2_not_in_set,
              S$answers[["post_cutoff_patients"]] - S$answers[["q3_not_in_ici_cohort"]],
              S$il2_same_day_multi))) |>
  kable(caption = "EMPI linkage between the RPDR extract and the ICI pull, joined on the character EMPI.")
```

# 5 Full pipeline code

```{r}
#| label: code-functions
#| eval: false
#| file: R/functions.R
```

```{r}
#| label: code-build
#| eval: false
#| file: R/build.R
```

```{r}
#| label: session
#| echo: false
#| output: asis
cat("\n---\n\n*Aggregate results only; cells below 11 suppressed. Built from `R/build.R` on ",
    format(S$built, "%Y-%m-%d"), ".*\n", sep = "")
```
