---
title: "Soluble IL-2R × ICI — Data Management"
subtitle: "Inputs, variable changes, result classification and linkage QC"
author: "Tianqi Ouyang"
date: "`r Sys.Date()`"
---

::: {.callout-note}
## About this page

All patient-level processing happens locally in `R/build.R` (run with
`Rscript R/build.R`). This page shows the code and the **aggregate** QC tables the
build writes to `cache/summary.rds`. Counts from 1 to 10 are shown as `<11`.
:::

```{r}
#| label: setup
#| include: false
suppressPackageStartupMessages({ library(data.table); library(knitr) })
source("R/functions.R")
have <- file.exists("cache/summary.rds")
if (have) S <- readRDS("cache/summary.rds")
```

```{r}
#| label: not-built
#| echo: false
#| output: asis
if (!have) cat("::: {.callout-warning}\n## Pipeline has not been run\n",
               "Run `Rscript R/build.R` locally, then re-render.\n:::\n")
```

```{r}
#| label: exit
#| include: false
if (!have) knitr::knit_exit()
```

# 1 Data sources

```{r}
#| label: tbl-sources
#| echo: false
data.table(
  Source = c("RPDR Mrn", "RPDR Lab", "RPDR Med", "RPDR Dia", "ICI data pull"),
  File   = c(sub("\\{.*", "Mrn.txt", S$files$rpdr), sub("\\{.*", "Lab.txt", S$files$rpdr),
             sub("\\{.*", "Med.txt", S$files$rpdr), sub("\\{.*", "Dia.txt", S$files$rpdr),
             S$files$ici),
  Use    = c("Patient set (one row per patient)",
             "Soluble IL2R results (plus creatinine, not used here)",
             "Independent check for ICI exposure",
             "Malignancy codes, for chart-review triage",
             "ICI administrations (Epic MAR 'New Bag')"),
  Rows   = c(fmt_n(S$flow$n[1]), fmt_n(S$lab_rows), fmt_n(S$med_rows), "-", fmt_n(S$ici_rows)),
  Patients = c(fmt_n(S$flow$n[1]), fmt_n(S$lab_patients), "-", "-", fmt_n(S$ici_patients))) |>
  kable(caption = "Inputs. The RPDR extract (IRB 2022P001091) was pulled 2026-08-11; the ICI pull is dated 2026-07-17. Neither is in this repository.")
```

# 2 ICI data pull

## 2.1 Variable changes

Two changes are made when the file is read, so every downstream object uses the
RPDR conventions:

| Raw | Cleaned | Why |
|---|---|---|
| `EMPI MRN` (column G) | `EMPI` | Same name as the RPDR key; joins need no renaming |
| `SCHEDULEDSTARTDTS` = `YYYY-MM-DDThh:mm:ssZ` | `SCHEDULEDSTARTDTS` = `YYYY-MM-DD` (Date) | ICI timing is compared at the day level |

```{r}
#| label: code-read-ici
#| eval: false
#| code: !expr src_of(c("read_ici", "parse_ici_date"))
```

```{r}
#| label: tbl-ici-format
#| echo: false
data.table(
  Check = c("`EMPI` values that are 9 digits (ICI pull)",
            "`EMPI` values that are 9 digits (RPDR Lab, IL2R rows)",
            "`SCHEDULEDSTARTDTS` values that failed to parse",
            "Earliest `SCHEDULEDSTARTDTS` (month)", "Latest `SCHEDULEDSTARTDTS` (month)"),
  Result = c(sprintf("%.1f%%", 100 * S$ici_empi_9digit),
             sprintf("%.1f%%", 100 * S$rpdr_empi_9digit),
             fmt_n(S$ici_date_na),
             format(S$ici_date_range[1], "%Y-%m"), format(S$ici_date_range[2], "%Y-%m"))) |>
  kable(caption = "Format checks. Both files carry EMPI as a 9-digit string, so they are joined as character without padding.")
```

## 2.2 Is the trailing "Z" a real UTC time?

If `SCHEDULEDSTARTDTS` were true UTC, a 09:00 Boston infusion would read 13:00 or
14:00 and the distribution below would peak in the afternoon. It peaks at 10:00–12:00
and falls away after 17:00 — ordinary clinic hours — so the clock time is local and
the `Z` is an export artefact. The calendar date is therefore taken **as written**,
with no time-zone shift.

```{r}
#| label: tbl-hour
#| echo: false
h <- copy(S$ici_hour)
h[, bin := fcase(hour <= 7, "00-07", hour >= 19, "19-23", default = sprintf("%02d", hour))]
h <- h[, .(N = sum(N)), by = bin]
h[, .(`Scheduled hour (as written)` = bin, Administrations = fmt_n(N),
      `%` = fmt_pct(N, sum(N)))] |>
  kable(caption = "Hour of SCHEDULEDSTARTDTS across all ICI administrations.")
```

## 2.3 What the file contains

```{r}
#| label: tbl-ici-action
#| echo: false
rbind(S$ici_mar_action[, .(Field = "MARACTIONDSC", Value = MARACTIONDSC, N)],
      S$ici_order_status[, .(Field = "ORDERSTATUSDSC", Value = ORDERSTATUSDSC, N)],
      S$ici_route[, .(Field = "ROUTEDSC", Value = ROUTEDSC, N)])[
  , Records := fmt_parts(N), by = Field][, .(Field, Value, Records)] |>
  kable(caption = "Every row is a MAR 'New Bag' (a dose hung), including the few whose order was later discontinued - so every row counts as ICI given.")
```

```{r}
#| label: tbl-ici-generic
#| echo: false
gg <- collapse_small(S$ici_generic[order(-patients), .(generic, N = patients)], "generic")
gg[, .(`ICI (generic)` = generic, Patients = fmt_n(N))] |>
  kable(caption = sprintf("%s distinct MEDICATIONDSC strings; every one names a known ICI by generic name, sponsor code or brand (%s unmatched). Each product is labelled by the first ICI or fixed-dose co-formulation it names; a patient appears under each label received.", fmt_n(S$ici_drug_desc_n), fmt_n(S$ici_unmatched_desc)))
```

```{r}
#| label: tbl-ici-blinded
#| echo: false
S$ici_blinded[, .(`Blinded or-placebo product` = fifelse(blinded, "Yes", "No"),
                  Descriptions = fmt_n(descriptions), Records = fmt_n(records),
                  Patients = fmt_n(patients))] |>
  kable(caption = sprintf("Blinded 'X OR PLACEBO' trial products do not establish ICI exposure, so they are left out of the ICI start date. %s patients in the whole pull have only blinded products; %s of them are in the post-2015 IL2R cohort.", fmt_n(S$ici_blinded_only_patients), fmt_n(S$groups[group == GROUP_LEVELS[3], sum(N)])))
```

```{r}
#| label: tbl-ici-year
#| echo: false
S$ici_by_year[, .(Year = year, Administrations = fmt_n(records), Patients = fmt_n(patients))] |>
  kable(caption = "ICI administrations by year. Coverage begins mid-2015 with Epic; 2026 is a partial year.")
```

# 3 RPDR soluble IL2R labs

## 3.1 Tests

```{r}
#| label: tbl-il2-tests
#| echo: false
S$il2_tests[, .(Group_Id, Test_Id, LOINC = Loinc_Code, Unit = unit,
                Records = fmt_parts(records), Patients = fmt_n(patients),
                Years = fifelse(records < SMALL_CELL | patients < SMALL_CELL, "--",
                        fifelse(first_year == last_year, as.character(first_year),
                                paste0(first_year, "-", last_year))))] |>
  kable(caption = "The four sIL2R test codes in the Lab file. Creatinine rows in the same file are not used.")
```

## 3.2 Result classification

```{r}
#| label: code-classify
#| eval: false
#| code: !expr src_of(c("NUM_REGEX", "CANCEL_REGEX", "classify_il2r_result"))
```

```{r}
#| label: tbl-il2-status
#| echo: false
S$il2_status[, .(`Result class` = as.character(status), Records = fmt_n(records),
                 Patients = fmt_n(patients), `Counts as a measurement` =
                   fifelse(status == STATUS_LEVELS[4], "No", "Yes"))] |>
  kable(caption = "Every IL2R row falls in one class. 'Numeric (Result_Text)' rows have a blank or placeholder Result with the value written in the report text (older mcsq assay).")
```

```{r}
#| label: tbl-il2-phrases
#| echo: false
p <- copy(S$il2_nonnum_phrases)
p[N < 3 & !grepl("^\\(", phrase), phrase := "(other short phrases)"]
p <- p[, .(N = sum(N)), by = .(status, phrase)]
p[, .(`Result class` = as.character(status), `Result string` = phrase, Records = fmt_n(N))] |>
  kable(caption = "The Result strings behind the two non-numeric classes (standard lab status text). Strings seen fewer than 3 times are grouped.")
```

```{r}
#| label: tbl-il2-values
#| echo: false
S$il2_value_summary[, .(Unit = unit, Values = fmt_n(n), Median = round(median),
                        IQR = paste0(round(p25), "-", round(p75)))] |>
  kable(caption = sprintf("Numeric IL2R values by unit (all dates). Numeric Result strings with an irregular shape: %s.", fmt_n(S$il2_numeric_irregular)))
```

## 3.3 Records by year

```{r}
#| label: tbl-il2-year
#| echo: false
y <- copy(S$il2_by_year)
y[, period := fcase(year <= 2014, "2005-2014", year %between% c(2018, 2020), "2018-2020",
                    default = as.character(year))]
y <- y[, .(records = sum(records)), by = period]
y[, .(`Draw year` = period, `IL2R records` = fmt_n(records))] |>
  kable(caption = "IL2R rows by year (all result classes). Results before 2015 are excluded from the cohort; the Epic-coded assay begins in 2021.")
```

# 4 Linkage

```{r}
#| label: tbl-linkage
#| echo: false
data.table(
  Check = c("IL2R patients (Lab) not in the RPDR patient set",
            "Post-2015 IL2R patients found in the ICI pull",
            "Patient-days with 2 or more performed IL2R results"),
  N = fmt_n(c(S$il2_not_in_set,
              S$answers[["post_cutoff_patients"]] - S$answers[["q3_not_in_ici_cohort"]],
              S$il2_same_day_multi))) |>
  kable(caption = "EMPI linkage between the RPDR extract and the ICI pull, joined on the character EMPI.")
```

# 5 Full pipeline code

```{r}
#| label: code-functions
#| eval: false
#| file: R/functions.R
```

```{r}
#| label: code-build
#| eval: false
#| file: R/build.R
```

```{r}
#| label: session
#| echo: false
#| output: asis
cat("\n---\n\n*Aggregate results only; cells below 11 suppressed. Built from `R/build.R` on ",
    format(S$built, "%Y-%m-%d"), ".*\n", sep = "")
```
